Identification of DC Pixels using Boosted Decision Trees

### 2016-05-11 report/bdtarticle.tex
Backup 11/5/16

Thus, once trained, the BDT was applied to every pixel entry in a 'testing' set of simulated telescope images. For each pixel, the BDT assigned a 'Signal Probability', $P_{signal}$, indicating the likelihood of the pixel being the DC pixel. From an entire image, the pixel with the largest $P_{signal}$ was identified as the DC pixel candidate. Repeating the double simulation technique for the test data, the true DC pixels in the test telescope images were reliably identified from EAS-free images. Thus, we can calculate the accuracy of BDT identification for test telescope images.

@@

In order to identify the DC pixel, a \textquoteleft DC Simulation' was initially run with an energy cut of 10 PeV on all muons and electrons. The consequence of this was to ensure that only the Cherenkov Light from the primary particle, as well as its fragments, was simulated. A second identical \textquoteleft Full Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete air shower including the DC light, representative of those imaged in telescope arrays. Comparisons of the Full and DC simulations enabled us to identify background light in the Full shower.

Using the sim\textunderscore telarray package \cite{Bernlohr08}, the expected hardware response to an air shower can be simulated. Again, the built-in response of the HESS telescope was selected for simulation. The program accounts for atmospheric transmission and density, mirror positions, sizes and reflectivities, camera shadowing and triggering, quantum efficiency and pulse responses. Due to the comprehensive and detailed nature of these hardware simulations, the resultant images can be considered representative of true camera images. For the Full Simulation, the night sky background was also simulated by sim\textunderscore telarray.

The HESS telescope has both a high gain Channel 0 and a low gain Channel 1. The location of each pixel (dependent on camera size), alongside the Sim\textunderscore telarray simulated value of SOMETHING? in each value was extracted. The pedestal and gain for each channel was also recorded, and from this, Intensity can be found.

@@

In the case of many iron core events, Channel 0 will reach its maximum value and become saturated. Thus Channel 0 ceases to be useful for discriminating between high energy DC and non-DC pixels, so was not used in this analysis. Instead, only the Channel 1 was used in later analysis. Thus the Channel 1 Intensity will simply be referred to throughout this analysis as $Intensity$.

In addition to the $Intensity$ of each pixel, the Sim\textunderscore telarray derives various Hillas whole-image parameters, which the script also extracted. These include the image width and length measured in degrees, from which the aspect ratio $A.R = \frac{width}{length}$ was calculated. Further the reconstructed shower direction and the shower center of gravity were calculated as positions in azimuth and zenith. Additionally the calculated energy and distance to core $r_{core}$ were recorded.

An entry was constructed for every pixel, containing its count and position. The variables $ \Delta_{C.o.G}$, $\Delta_{Direction}$ and $\Delta_{Line}$ were defined as the distance from the pixel to the shower center of gravity, shower direction, and the line joining those two points. Furthermore, the nearest neigbouring pixel IDs were calculated for every pixel position, enabling the count in each neighbouring pixel to be found. From this, two further pixel variables were defined. The largest neighbouring intensity $Intensity_{N.N.max}$ was found, and the ratio $ Q_{DC} = \frac{Intensity}{Intensity_{N.N.max}} $ was derived. In addition the Nearest Neighbour Mean Count was recorded. These variables were recorded in every pixel entry, and together the pixel entries formed a complete dataset for each camera image.

\subsection{Classic DC Identification}
The original method used by the HESS collaboration \cite{hess07} was initially replicated, for which a number of cuts were applied to each image dataset.

\begin{table}[h!]
  \centering
  \caption{Cuts applied to image pixel sets}
  \label{tab:table1}
  \begin{tabular}{ccc}
    \toprule
    Variable & Cut\\
    \midrule
     $r_{core}$ & \textgreater 0.4 \\
     $Q_{DC}$ & \textgreater 1.3 \\
     $ \Delta_{C.o.G}$ & \textgreater 0.17 \\
     $ \Delta_{C.o.G}$ & \textless 0.91 \\
     $\Delta_{Direction}$ & \textless 0.45 \\
     $\Delta_{Line}$ & \textless 0.23 \\
     Aspect Ratio & \textless 0.75 \\
    \bottomrule
  \end{tabular}
\end{table}

From the subset of pixels satisfying these conditions, the remaining pixel with the largest $Q_{DC}$ was selected as the DC pixel. The result was compared with the true DC pixel, identified from the hadron-only image, to test the accuracy of this cut. Out of 400 events, it was fount that only 55\% of these events contained a clear DC signal, as defined by requiring the DC pixel to have $Intensity_{DC} > 150$ in the hadron-only image. SOMETHING! Of this subset of 'DC events', 17\% events passed all of the required cuts. The $Q_{DC}$ was found to be 100\% accurate in identifying the DC pixel in those passing events, as shown in \ref{fig:cutdistribution}.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{cutdistributionhess1None}
\caption{The DC signal in the shower-free pixel is shown in the top left, with a broad gaussian distribution with the tail of the night-sky background extending up to appproximately 5000. Events below this are unlikely to be identified correctly because the DEC light is too faint. In the top right- the distribution of the dataset is shown, once all the non-$Q_{DC}$ cuts have been applied. In the bottom left, the BDT score distribution is shown before any cuts. On the bottom right, we see the same distribution after both signal and BDT score cuts are applied. All green events are ones in which the DC pixel has been correctly identified, while red events are ones that have been incorrectly identified.}
\label{fig:cutdistribution}
\end{center}
\end{figure}

\subsection{Boosted Decision Tree DC Identification}  As an alternative to use of $Q_{DC}$, a new method of DC pixel identification was developed as part of this analysis by using a BDT trained with the Scikit Learn Python package. A set of 2000 simulated events was selected, and randomly split through use of the python random.random() function into two subsets for training and testing. The variable $DC_{Count} = Count-Mean_{N.N}$ was defined as an approximate \textquoteleft DC signal', if the pixel were to be selected as the DC candidate. For every pixel in the 2000 simulated events HESS 1 images, an entry was formed of $Q_{DC}$, $ \Delta_{C.o.G}$, $\Delta_{Direction}$, $\Delta_{Line}$, Channel1, Channel0, $DC_{Count}$ and $Mean_{N.N}$. A score of 0 was assigned to every pixel to indicate background, with the exception of those that were identified as the DC pixel through the Shower-Free simulation. If the DC count in the shower-free simulation was larger than 300, the pixel was included with a score of 1, identifying it as a signal pixel. Pure DC pixels below the threshold cut of 300 were deemed to be too faint to reconstruct, so these entries were omitted from the training set.

@@

\subsection{Composite Event Selection}
Having trained the BDT successfully, it was then applied to the same dataset as for the classic QDC identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $Probability_{signal} > 0.5$ for the DC candidate to be accepted.

@@

\subsection{Error in calculated $Count_{DC}$}
Having successfully identified the DC pixel in many events, we can calculate the difference between the derived $candidate_{DC count}$ and the hadron-only $True_{DC}$ to determine the error we will find for the LPD. Binning the difference $\Delta = candidate_{DC count} - True_{DC count}$ as shown in FIG, we find that the mean difference is 65 and the median difference is -75. We thus see that the distribution is not significantly skewed. However, the Standard deviation is approximately 600 when measured both by half of the difference from the 16th to 84th centile, and by 68th centile of an absolute difference. This again suggests that, although skewed, the fraction error in intensity is rather large. However, the fractional error found purely on the basis of QDC calculations is EQUALLY LARGE??LARGER?.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{DCcounterrorhess1}
\caption{The difference between true and reconstructed DC signal for events passing the cuts is shown in the top left. In the top right the same is shown for events which were rejected. The true DC signals for events which passed the cuts is plotted against the reconstructed signals in the lower left, and for rejected events in the lower right. As in \ref{fig:cutdistribution}, in all plots a green event is one in which the DC pixel has been correctly identified, while a red event is one that has been incorrectly identified.}
\label{fig:dcdiff}
\end{center}
\end{figure}

>>>>

Once trained, the BDT was applied to every pixel entry in a separate 'testing' set of simulated telescope images. For each pixel, the BDT assigned a \textquoteleft Signal Probability', $P_{signal}$, indicating the likelihood of the pixel being the DC pixel. From an entire image, the pixel with the largest $P_{signal}$ was identified as the DC pixel candidate. Repeating the double simulation technique for the test data, the true DC pixels in the test telescope images were reliably identified from EAS-free images. Thus, the accuracy of BDT identification for test telescope images was calculated.

@@

In order to identify the DC pixel, a simulation was initially run with an energy cut of 10 PeV on all muons and electrons. As the primary particle energy did not exceed this value, neither daughter muons and electrons, nor the photons they would have emitted, were simulated. Thus the hadronic Cherenkov Light from the primary particle and daughter fragments, but not the EAS light, was present in the camera image. A second identical \textquoteleft EAS Simulation' was run including the same random seeds, but without the energy cut on muons and electrons. This gave a complete EAS image including the DC light.

Using the sim\textunderscore telarray package \cite{Bernlohr08}, the expected HESS hardware response to each air shower was simulated. Among other things, the program accounts for atmospheric transmission and density, mirror positions, sizes and reflectivities, camera shadowing and triggering, quantum efficiency and pulse responses. Due to the comprehensive and detailed nature of these hardware simulations, the resultant images can be considered \textquoteleft realistic camera images'. For the EAS Simulation, the night sky background was also included by sim\textunderscore telarray, further obscuring the DC pixel.

The HESS telescope has both a high gain Channel 0 and a low gain Channel 1. The Sim\textunderscore telarray simulated value of SOMETHING? for each channel were recorded. The pedestal and gain for each channel were also recorded, and from this, Intensity in the pixel channel was found.

@@

In the case of many iron core events, Channel 0 will reach its maximum value and become saturated. Thus Channel 0 ceases to be useful for discriminating between high energy DC and non-DC pixels. Consequently, Channel 0 Intensities neglected while Channel 1 Intensities were used in all later analysis. This variable will henceforth be referred to simply as $Intensity$.

In addition to the $Intensity$ of each pixel, the Sim\textunderscore telarray package derives various Hillas whole-image parameters. These include the image width and length measured in degrees, from which the aspect ratio $A.R = \frac{width}{length}$ was calculated. Further, the reconstructed shower direction and the shower center of gravity were calculated as positions in azimuth and zenith. Additionally the calculated energy and distance to core $r_{core}$ were recorded.

For every pixel, the location within the image was determined from a standard HESS layout. The variables $ \Delta_{C.o.G}$, $\Delta_{Direction}$ and $\Delta_{Line}$ were defined as the distance from the pixel to the shower center of gravity, shower direction, and the line joining those two points. Furthermore, the nearest neigbouring pixel IDs were calculated for every pixel position, enabling the count in each neighbouring pixel to be found. From this, two further pixel variables were defined. The largest neighbouring intensity $Intensity_{N.N.max}$ was found, and the ratio $ Q_{DC} = \frac{Intensity}{Intensity_{N.N.max}} $ was derived. In addition the Nearest Neighbour Mean Count was recorded. These variables were recorded in every pixel entry, and together the pixel entries formed a complete dataset for each camera image.

\section{DC Pixel Identification}  
The original method used by the HESS collaboration \cite{hess07} was initially replicated, for which a number of cuts were applied to each image dataset.

\begin{table}[h!]
  \centering
  \caption{Cuts applied to image pixel sets}
  \label{tab:table1}
  \begin{tabular}{ccc}
    \toprule
    Variable & Cut\\
    \midrule
     $Q_{DC}$ & \textgreater 1.3 \\
     $ \Delta_{C.o.G}$ & \textgreater 0.17 \\
     $ \Delta_{C.o.G}$ & \textless 0.91 \\
     $\Delta_{Direction}$ & \textless 0.45 \\
     $\Delta_{Line}$ & \textless 0.23 \\
     Aspect Ratio & \textless 0.75 \\
    \bottomrule
  \end{tabular}
\end{table}

From the subset of pixels satisfying these conditions, the remaining pixel with the largest $Q_{DC}$ was selected as the DC pixel. In the original analysis, an additional cut $r_{core} \textgreater 0.4$ was applied. However, the uncertainty in determining the core position through Hillas Analysis is typically of the order of $\pm 30m$. Consequently, this particular cut was omitted, along with the Impact Parameter cuts. For every image, the total image amplitude $I_{tot}$ was found. A cut was also placed requiring $Q_{DC} < 0.14 \times \log(\dfrac{I_{tot}}{(161 \times \cos \theta)})$ where $\theta$ was the zenith angle. The result was compared with the true DC pixel, identified from the hadron-only image, to test the accuracy of this cut. Out of 400 events, it was fount that only 55\% of these events contained a clear DC signal, as defined by requiring the DC pixel to have $Intensity_{DC} > 150$ in the hadron-only image. SOMETHING! Of this subset of 'DC events', 17\% events passed all of the required cuts. The $Q_{DC}$ was found to be 100\% accurate in identifying the DC pixel in those passing events, as shown in \ref{fig:cutdistribution}.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{cutdistribution1None}
\caption{The DC signal in the shower-free pixel is shown in the top left, with a broad gaussian distribution with the tail of the night-sky background extending up to appproximately 5000. Events below this are unlikely to be identified correctly because the DEC light is too faint. In the top right- the distribution of the dataset is shown, once all the non-$Q_{DC}$ cuts have been applied. In the bottom left, the BDT score distribution is shown before any cuts. On the bottom right, we see the same distribution after both signal and BDT score cuts are applied. All green events are ones in which the DC pixel has been correctly identified, while red events are ones that have been incorrectly identified.}
\label{fig:cutdistribution}
\end{center}
\end{figure}

As an alternative to use of $Q_{DC}$, a new method of DC pixel identification was developed as part of this analysis by using a BDT trained with the Scikit Learn Python package. A set of 2000 simulated events was selected, and randomly split through use of the python random.random() function into two subsets for training and testing. The variable $DC_{Count} = Count-Mean_{N.N}$ was defined as an approximate \textquoteleft DC signal', if the pixel were to be selected as the DC candidate. For every pixel in the 2000 simulated events HESS 1 images, an entry was formed of $Q_{DC}$, $ \Delta_{C.o.G}$, $\Delta_{Direction}$, $\Delta_{Line}$, Channel1, Channel0, $DC_{Count}$ and $Mean_{N.N}$. A score of 0 was assigned to every pixel to indicate background, with the exception of those that were identified as the DC pixel through the Shower-Free simulation. If the DC count in the shower-free simulation was larger than 300, the pixel was included with a score of 1, identifying it as a signal pixel. Pure DC pixels below the threshold cut of 300 were deemed to be too faint to reconstruct, so these entries were omitted from the training set.

@@

Having trained the BDT successfully, it was then applied to the same dataset as for the classic QDC identification. In each camera image, the event with the largest BDT score was deemed to be \textquoteleft most signal-like', and thus selected as the DC pixel candidate. A cut was applied, requiring $Probability_{signal} > 0.5$ for the DC candidate to be accepted.

@@

\section{Conclusion}
Basically it just worked really well.
